
---
output: 
  html_document:
  pdf_document: default
  word_document: default
title: "Assignment 7 - Extra Credits: Practice dplyr and ggplot2"
---

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](assignment7_extra_credits.Rmd)) in Rstudio. 

- Right under each **question**, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Canvas



```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```


-------

1. Import the data at https://bryantstats.github.io/math421/data/ds_salaries.csv

    The data description can be found at https://www.kaggle.com/datasets/saurabhshahane/data-science-jobs-salaries
```{r}

library(tidyverse)
setwd("C:/Users/student/Downloads")
df =  read_csv('ds_salaries.csv')
head(df)


```

2. Is there any column that contains NAs?
```{r}

colSums(is.na(df))


```

3. What is the most frequent job title in the data?  
```{r}

df %>% count(job_title) %>% 
  arrange(-n)
  


```

4. Find the average salary data of the three most frequent job titles in the data. 
```{r}

df %>% 
  filter(job_title=='Data Scientist'|job_title=='Data Engineer'|job_title=='Data Analyst') %>%
  group_by(job_title) %>% 
  summarise(mean(salary))
  


```

5. What is the job title that have the highest average salary?
```{r}

df %>% 
  group_by(job_title) %>% 
  summarise(meansal = mean(salary)) %>% 
  arrange(-meansal)


```

6. Which country gives the highest salary on average? 
```{r}

df %>% 
  group_by(company_location) %>% 
  summarise(meansal = mean(salary)) %>% 
  arrange(-meansal)


```

7. Plot five meaningful plots. All plots should have title, caption, and appropriate labels.
```{r}

colSums(is.na(df))
df %>% 
  group_by(employment_type, work_year) %>% 
  summarise(meanSal = mean(salary_in_usd)) %>% 
  ggplot()+
  geom_col(aes(x=work_year, y=meanSal, fill=employment_type), position = 'dodge')+ labs(
       title = 'Mean salary by work year and employment type', 
       caption = 'Employment type CT went way up in mean salary in 2021, but then decreased very much the next year')


```
```{r}

df %>% 
    filter(company_size=='L') %>% 
    ggplot()+
    geom_bar(mapping=aes(x=experience_level))+ labs(
       title = 'Number of jobs by experience level that are large companies', 
       caption = 'The experience levels MI and SE and the most common in large companies')


```
```{r}

df %>% ggplot()+ 
  geom_bar(mapping=aes(x=work_year, fill=experience_level), 
           position = 'fill')+theme_dark()+ labs(
       title = 'Percentage of jobs in each year that are each experience level ', 
       caption = 'In 2022, the experience level SE went way up and was the highest percentage job that year')


```

```{r}


df %>% ggplot()+ 
  geom_bar(mapping=aes(x=company_size, fill=salary_currency), 
           position = 'dodge')+ labs(
       title = 'Number of jobs by company size and salary_currency', 
       caption = 'USD is the most common currency in each size of company')


```
```{r}
df %>% 
  group_by(remote_ratio, company_size) %>% 
  summarise(meanSal = mean(salary_in_usd)) %>% 
  ggplot()+
  geom_col(aes(x=remote_ratio, y=meanSal, fill=company_size), position = 'dodge')+ labs(
       title = 'Mean salary by remote work ratio and company size', 
       caption = '100% remote salary is consistently the highest, and for 50% remote, the salary is higher for medium-sized companies than large ones')


```

