
---
output: # rmarkdown::github_document
  html_document:
  pdf_document: default
  word_document: default
title: "Assignment 3.  Working with Data Frame. Base R Style"
---

***Note*:** *This assignment practices working with Data Frame using Base R.*

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](assignment3.Rmd)) in Rstudio. 

- Right under each question, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Canvas under Assignment 3. 

-------

Problems
-------
 \

1. Create the following data frame
```{r}
df = data.frame(Rank = c(0,1,2,3),
                Age = c(28,34,29,42),Name = c('Tom','Jack','Steve','Ricky'))

df
```

| Rank 	| Age 	| Name  	|
|:------	|:-----	|:-------	|
| 0    	| 28  	| Tom   	|
| 1    	| 34  	| Jack  	|
| 2    	| 29  	| Steve 	|
| 3    	| 42  	| Ricky 	|

2. Use `read.csv` to import the Covid19 Vaccination data from WHO: [link](https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-states.csv).
```{r}
df = read.csv('https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-states.csv')
head(df)
```

2. Show the names of the variables in the data
```{r}
names(df)
```

3. How many columns and rows the data have? 
```{r}
str(df)

# 5 columns with 61,942 rows
```

4. How many missing values are there?  Show the missing values by columns.  What variable has the most number of missing values?
```{r}
sum(is.na(df))
colSums(is.na(df))
colMeans(is.na(df))
```

5. What is the class of the `date` column. Change the `date` columns to `date` type using the `as.Date` function. Show the new class of the `date` column. 
```{r}
class(df$date)
df$date = as.Date(df$date)
class(df$date)

```

6. Capitalize the names of all the variables
```{r}

names(df)[1:5] <-  c("Date" ,"State" , "Fips" ,"Cases" , "Deaths")
names(df)
```

7. Find the average number of cases per day. Find the maximum cases a day.  
```{r}
mean(df$Cases, na.rm=TRUE)

max(df$Cases, na.rm=TRUE)
```

8. How many states are there in the data?  
```{r}

length(table(df$State))

```

9. Create a new variable `weekdays` to store the weekday for each rows. 
```{r}
df$weekdays <- weekdays(df$Date)
head(df)
```

10. Create the categorical variable `death2` variable taking the values as follows

- `has_death` if there is a death that day
-  `no_death` if there is no death that day
```{r}
df$death2 <- ifelse(df$Deaths==0, 
                        'no_death','has_death') 
table(df$death2)
```

Find the frequency and relative frequency of `no_death` and `has_death`. 
```{r}
table(df$death2)
prop.table(table(df$death2))

```

11. Find the first quartile (Q1), second quartile (Q2) and and third quartile (Q3) of the variable `death`. (Hint:  Use the `summary` function)
```{r}
summary(df$Deaths)
```

12. Create the categorical variable `death3` variable taking the values as follows

- `low_death` if the number of deaths smaller than the 25 percentile (Q1)

- `mid_death` if the number of deaths from Q1 to Q3

- `high_death` if the number of deaths greater than Q3
```{r}
library(dplyr)
df$death3 = case_when(df$Deaths <1171 ~ 'low_death',
                             df$Deaths<14912 ~ 'mid_death',
                             TRUE~'high_death')
table(df$death3)
```

13. Find the average cases in Rhode Island in 2021
```{r}
df4 <- df[(df$State=='Rhode Island')&
            (df$Date>='2021-01-01')&
            (df$Date<='2022-01-01'),]

mean(df4$Cases, na.rm=TRUE)
```

14. Find the median cases by weekdays in Rhode Island in 2021 
```{r}

by(df4$Cases, df4$weekdays, median)
```

15. Compare the median cases in Rhode Island in June, July, August and September in 2021. 
```{r}
library(lubridate)
library(dplyr)
df4$month <- month(df4$Date, label = TRUE)
df4$month2 <- case_when(df4$month %in% c('Oct','Nov','Dec') ~ 'fall_semester',
                       df4$month %in% c('Jan','Feb','Mar','Apr','May') ~ 'spring_semester',
                       TRUE~'break')
df5 <- df4[df4$month2=='break',]
by(df5$Cases, df5$month, median)

```

16.  Find your own dataset, import it and implement the following functions on the data
```{r}
setwd("C:/Users/student/Downloads")
df <-  read.csv('titanic.csv')
head(df)
```
```{r}

head(df)


```


```{r}

names(df)

```


```{r}

str(df)

```


```{r}

mean(df$Age, na.rm=TRUE)
min(df$Age, na.rm=TRUE)
max(df$Age, na.rm=TRUE)


```


```{r}
table(df$Pclass)
```


```{r}

sum(is.na(df))
colSums(is.na(df))

```


```{r}

class(df$Embarked)


```


```{r}

cor(df$Fare, df$Survived)

```


```{r}

by(df$Survived, df$Pclass, FUN = mean)

```


```{r}

df$survival <- ifelse(df$Survived==0, 
                        'Survived','Died') 
table(df$survival)

```
```{r}

library(dplyr)
df$Survival2 <- case_when(df$Survived==0 ~ 'Survived', 
                            TRUE ~ 'Died') 
table(df$Survival2)

```

- head
- str
- names
- mean, min, max
- table
- is.na
- colSums
- class
- cor
- by
- ifelse
- case_when

17.  In the dataset in #16, practice the follows.  You can reuse the code of 16.


- Create a categorical variable from a continuous variable
```{r}

df$Age2 <- case_when(df$Age<20 ~ 'low_age',
                             df$Age<50 ~ 'mid_age',
                             TRUE~'high_age')
table(df$Age2)


```

- From one categorical variable, create a new categorical variable with fewer categories
```{r}


df$Siblings <- case_when(df$SibSp %in% c('0','1') ~ 'few_sibs',
                       df$SibSp %in% c('2','3','4','5','6','7','8') ~ 'manySibs',
                       TRUE~'break')
table(df$Siblings)

```



If you do not have a data, you can use [the titanic dataset](https://www.kaggle.com/competitions/titanic/overview), which can be downloaded at [this link](../data/titanic.csv)
