
---
output: 
  html_document:
  pdf_document: default
  word_document: default
title: "Assignment 7: Visualization - Some aspects of ggplot2"
---

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](assignment7.Rmd)) in Rstudio. 

- Right under each **question**, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Canvas

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

-------

Use the data of your own.  Produce the following types of plots and comment on each plot. Plots should be meaningful. If you use the data we used in class, make sure the plots are not the same as the ones in the slides. All plots should have title, caption, appropriate labels on x and y-axis. 

-------
```{r}

library(tidyverse)
setwd("C:/Users/student/Downloads")
df =  read_csv('Spotify-2023.csv')
head(df)

```

1. Make a plot using `facet_wrap` by one categorical variable (`facet_wrap(~variable1)`)
```{r}

df %>% ggplot()+
  geom_point(aes(x=in_apple_charts, y=in_spotify_charts))+facet_wrap(~mode)+
  labs(
       title = 'Number of Apple vs Spotify Charts by Mode', 
       caption = 'Major songs tend to have a similar proportion of apple to spotify, while Minor songs are not as correlated')

## songs on apple vs spotify charts by mode
```

2. Make a plot using `facet_wrap` by two categorical variables (`facet_wrap(~variable1+variable2)`)
```{r}

df %>% ggplot()+
  geom_point(aes(x=in_apple_charts, y=in_spotify_charts))+facet_wrap(~mode+key)+labs(
       title = 'Spotify vs Apple charts by Mode and Key', 
       caption = 'Minor F key and Major C# key have many songs high on the apple charts ')

## songs on apple vs spotify charts by mode and number of artists per song
```

3. Make a plot using `position ='dodge'`.
```{r}

df %>% ggplot()+ 
  geom_bar(mapping=aes(x=mode, fill=key), 
           position = 'dodge')+ labs(
       title = 'Number of songs by Mode and Key', 
       caption = 'Major in general has more songs, except for D# and E Major songs')


```

4. Make a plot using `position ='fill'`. 
```{r}

df %>% ggplot()+ 
  geom_bar(mapping=aes(x=mode, fill=key), 
           position = 'fill')+ labs(
       title = 'Percentage of songs by Mode and Key', 
       caption = 'Major songs have a lot of C#, D, G and G#, while Minor songs are much more evenly spread')


```

5. Make a plot using `geom_col`.
```{r}

df %>% group_by(key) %>% 
  summarise(average_apple=mean(in_apple_playlists, na.rm=TRUE)) %>% 
  ggplot()+ 
  geom_col(mapping=aes(x=key, y=average_apple))+ labs(
       title = 'Number of songs on apple playlists by Key', 
       caption = 'C# songs have the highest number of songs on apple playlists')


```

6. Change the theme of one of your plots. 
```{r}

df %>% ggplot()+ 
  geom_bar(mapping=aes(x=mode, fill=key), 
           position = 'fill')+theme_dark()+ labs(
       title = 'Percentage of songs by Mode and Key', 
       caption = 'Minor songs are much more evenly spread out among the different keys than for Mode')

```

7. Make a plot using combo filter + ggplot
```{r}

df %>% 
    filter(artist_count>3) %>% 
    ggplot()+
    geom_bar(mapping=aes(x=mode, fill=key))+ labs(
       title = 'Number of songs by Mode and Key which have 3 or more artists collaborating on the song', 
       caption = 'The number of Major songs is smaller than Minor, and G# makes up the largest percentage of Minor songs')


```

8. Make a plot using combo group_by + summarise + geom_col
```{r}

df %>% 
  group_by(mode, key) %>% 
  summarise(mean_artists = artist_count) %>% 
  ggplot()+
  geom_col(aes(x=mode, y=mean_artists, fill=key), position = 'dodge')+ labs(
       title = 'Average number of artists by Mode and Key', 
       caption = 'The highest average number of artists for Minor songs is for G key songs')


```

9. Save a plot and put a link of the plot to your Github Webpage.  To link a plot `abc.png` you just need to insert the following on the README.md

```{r, eval=FALSE}
[Link to the plot](abc.png)
```


```{r, eval=FALSE}
gg <- df %>% group_by(mode, key) %>% summarise(mean_artists=mean(artist_count)) %>% 
  ggplot()+ 
  geom_col(mapping=aes(x=mode, y=mean_artists, fill=key), position = 'dodge')
ggsave(filename = 'abc.png', plot = gg)

```

10. In this question, we will use  the Covid19 vaccination data for the US. The data and data dictionary can be found at the below link:

- [Data](https://raw.githubusercontent.com/owid/covid-19-data/master/public/data/vaccinations/vaccinations-by-manufacturer.csv)

- [Alternative Link](https://bryantstats.github.io/math421/data/vaccinations-by-manufacturer.csv)

```{r}
library(tidyverse)
df <- read_csv('https://raw.githubusercontent.com/owid/covid-19-data/master/public/data/vaccinations/vaccinations-by-manufacturer.csv')
head(df)
```



`Geom_point` makes scatter plots, which plot two numeric (or Date) variables. A third categorical variable is often brought in for color. For example, we can compare different vaccine types used in the US by

```{r}
df %>% 
  filter(location=='United States') %>% 
  ggplot()+
  geom_point(mapping = aes(x = date, 
                                    y = total_vaccinations, 
                                    color = vaccine))+ labs(
       title = 'Total Vaccines in the US by date', 
       caption = 'The number of vaccines increase greatly, then level off starting around 2022')
```

Do the follows. 

  - Use `geom_point` to compare the vaccines used by date in Argentina in 2022. 
```{r}
df$year = year(df$date)
df %>% 
  filter(location=='Argentina',year==2022) %>% 
  ggplot()+
  geom_point(mapping = aes(x = date, 
                                    y = total_vaccinations, 
                                    color = vaccine)) + labs(
       title = 'Total Vaccines in Argentina in 2022 by Date', 
       caption = 'The most common vaccine throughout is Sinopharm/Beijing')


```
  
  - Change `geom_point` to `geom_line` in the above codes to have the corresponding line plots. 
```{r}

df %>% 
  filter(location=='Argentina',year==2022) %>% 
  ggplot()+
  geom_line(mapping = aes(x = date, 
                                    y = total_vaccinations, 
                                    color = vaccine)) + labs(
       title = 'Total Vaccines in Argentina in 2022 by Date', 
       caption = 'The most common vaccine throughout is Sinopharm/Beijing')


```

11. (Using the same dataset as in 10.) Sometime we need to do some calculations on the data before plotting. For example, the below codes is to compare the total vaccines (including all types of vaccines) by date between the US and the EU

```{r}
# the total vaccines (including all types of vaccines) by date
d1 <- df %>% 
  filter(location %in% c('United States', 'European Union')) %>% 
  group_by(date, location) %>% 
  summarise(total_vac2=sum(total_vaccinations, na.rm=TRUE))

# Plot
d1 %>% 
  ggplot()+
  geom_point(mapping=aes(x=date, 
                         y=total_vac2, 
                         color=location)) + labs(
       title = 'Total Vaccines in the US and EU by Date', 
       caption = 'The EU greatly outnumbers the US starting midway through 2021')
```

Notice that the above codes could be piped in one 

```{r, eval=FALSE}
df %>% 
  filter(location %in% c('United States', 'European Union')) %>% 
  group_by(date, location) %>% 
  summarise(total_vac2=sum(total_vaccinations, na.rm=TRUE)) %>% 
  ggplot()+
  geom_point(mapping=aes(x=date, 
                         y=total_vac2, 
                         color=location))+ labs(
       title = 'Total Vaccines in the US and EU by Date', 
       caption = 'The EU greatly outnumbers the US starting midway through 2021')
```


Do the follows. 

- Use `geom_point` compare the total vaccines (including all types of vaccines) by date between countries/regions in Asia
```{r}
df %>% 
  filter(location %in% c('Hong Kong', 'Japan','South Korea','Nepal')) %>% 
  group_by(date, location) %>% 
  summarise(total_vac2=sum(total_vaccinations, na.rm=TRUE)) %>% 
  ggplot()+
  geom_point(mapping=aes(x=date, 
                         y=total_vac2, 
                         color=location)) + labs(
       title = 'Total Vaccines in Asia by Date', 
       caption = 'Japan outnumbers all other countries by a lot starting summer 2021')


```

- Use `geom_point` compare the total vaccines (including all types of vaccines) by date between countries/regions in South America in 2021
```{r}

df %>% 
  filter(location %in% c('Argentina', 'Chile','Ecuador','Peru'), year==2021) %>% 
  group_by(date, location) %>% 
  summarise(total_vac2=sum(total_vaccinations, na.rm=TRUE)) %>% 
  ggplot()+
  geom_point(mapping=aes(x=date, 
                         y=total_vac2, 
                         color=location)) + labs(
       title = 'Total Vaccines in South America by Date', 
       caption = 'Argentina starts outnumbering the other countries starting in July 2021')


```

