
---
output: # rmarkdown::github_document
  html_document:
    theme: cerulean
    toc: true
  pdf_document: default
  word_document: default
title: "Assignment 5: Handle Missing Values"
---

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](assignment5.Rmd)) in Rstudio. 

- Right under each **question**, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Canvas. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

-------

**The data:** 

This assignment works with the Adult Census Data that can be downloaded at [this link.](https://bryantstats.github.io/math421/data/adult_census_missing.csv)

-------

**Questions**

1. Use `read_csv` to import the data.  Show the number of NAs for each columns. 
```{r}

library(tidyverse)
setwd("C:/Users/student/Downloads")
df =  read_csv('adult_census_missing.csv')
colSums(is.na(df))

```


2. Using function `aggr` the VIM package to plot the number of NAs for each column. 
```{r}
#install.packages("VIM")
library(VIM)
aggr(df)

```

3. Find other forms of missing values. Hint: You can use the `table` function to check if there are suspicious categories (Unknown, for example) in a variable. What all other forms of missing values shown in the data. 
```{r}

colSums(df=='Unknown', na.rm = TRUE)

colSums(df=='?', na.rm = TRUE)

colSums(df=='Not Known', na.rm = TRUE)


```

4. Replace all the forms of missing values found to NA
```{r}

df <- replace(df, df == 'Unknown' |
                  df == 'Not Known' |
                  df == '?', NA)
head(df)

```

5. Replot the number of NAs for each column. 
```{r}

aggr(df)



```

6. Approach 1 to handle NAs: remove all rows that have any NAs. Save the dataset after removing as a different data. The original data is unchanged (still have NAs).  How many rows left after removing?
```{r}

df2 = drop_na(df)
glimpse(df2)

## 29,865 rows

```

7. Approach 2 to handle NAs:  Fill all the NAs with the previous or next value. (Hint: Using `fill` function).  Save the dataset after filling as a different data. The original data is unchanged (still have NAs). 
```{r}

df <- replace(df, df == 'Unknown' |
                  df == 'Not Known' |
                  df == '?', NA)
df3 = df %>% fill( age, workclass, fnlwgt, education, education.num, marital.status , occupation  , relationship, race, sex, capital.gain, capital.loss, hours.per.week, native.country, income, .direction = 'updown')
head(df3)


```

8. Approach 3 to handle NAs: For numeric variable, replace the NAs by the median.  For categorical variables, replace the NAs to the majority. 
```{r}

med_age <- median(df$age, na.rm = TRUE)
med_fnlwgt <- median(df$fnlwgt, na.rm=TRUE)
med_education.num <- median(df$education.num, na.rm=TRUE)
med_capital.gain <- median(df$capital.gain, na.rm=TRUE)
med_capital.loss <- median(df$capital.loss, na.rm=TRUE)
med_hours.per.week <- median(df$hours.per.week, na.rm=TRUE)
med_age
majority_workclass <- names(which.max(table(df$workclass)))
majority_education <- names(which.max(table(df$education)))
majority_occupation <- names(which.max(table(df$occupation)))
majority_relationship <- names(which.max(table(df$relationship)))
majority_native.country <- names(which.max(table(df$native.country)))
majority_sex <- names(which.max(table(df$sex)))



df$workclass <- replace_na(df$workclass, majority_workclass)
df$education <- replace_na(df$education, majority_education)
df$occupation <- replace_na(df$occupation, majority_occupation)
df$relationship <- replace_na(df$relationship, majority_relationship)
df$native.country <- replace_na(df$native.country, majority_native.country)
df$sex <- replace_na(df$sex, majority_sex)


df$age <- replace_na(df$age, med_age)
df$fnlwgt <- replace_na(df$fnlwgt, med_fnlwgt)
df$education.num <- replace_na(df$education.num, med_education.num)
df$capital.gain <- replace_na(df$capital.gain, med_capital.gain)
df$capital.loss <- replace_na(df$capital.loss, med_capital.loss)
df$hours.per.week <- replace_na(df$hours.per.week, med_hours.per.week)
head(df)



```





