
---
output: # rmarkdown::github_document
  html_document:
  pdf_document: default
  word_document: default
title: "Assignment 4 - Extra Credits"
---

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](assignment4_extra_credits.Rmd)) in Rstudio. 

- Right under each **question**, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- Notice that if there is `eval=FALSE` in the first line of the code chunk, the chunk will not be execute.

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Canvas. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

-------

This assignment works with the [IMDB Top 1000 data](../data/imdb_top_1000.csv).  Find out more information about this data at [this link](https://www.kaggle.com/datasets/harshitshankhdhar/imdb-dataset-of-top-1000-movies-and-tv-shows).  Import the data and answer the following questions. 

1. List all the names of the columns of the data
```{r}
  library(tidyverse)
setwd("C:/Users/student/Downloads")
df =  read_csv('imdb_top_1000.csv')
names(df)
```

2. Which movies have the highest money earned (Gross)?
```{r}



df %>% 
  select(Series_Title,Gross) %>% 
  arrange(-Gross)


```

3. What is the lowest rating (IMDB_Rating)?  List five movies have this lowest rating.
```{r}

tail(df,5) %>% 
  select(Series_Title,IMDB_Rating) %>% 
  arrange(IMDB_Rating) 



```

4. Which year have the most number of movies released in the list?  What is the total of money earned on that year?
```{r}

df %>% 
  select(Released_Year,Gross,Series_Title) %>% 
  group_by(Released_Year, Series_Title) %>% 
  summarise(Gross = sum(Gross)) %>% 
  count(Series_Title) %>% 
  summarise(n = sum(n)) %>% 
  arrange(-n)

```

5. What is the total money earned per movies on average?
```{r}
df %>% 
  summarise(money = mean(Gross, na.rm = TRUE))

  
```

6. Calculate the average number of votes by year. Calculate the average number of votes of movies that have IMDB rating greater than 9.
```{r}


df %>% 
  group_by(Released_Year) %>% 
  summarise(Votes = mean(No_of_Votes)) %>% 
  arrange(-Votes)


```
```{r}

df %>% 
  filter(IMDB_Rating>9) %>% 
  summarise(Votes = mean(No_of_Votes))  

```

7. Calculate the average Meta score in 2020 of movies that have number of votes in the third quartile. 
```{r}

summary(df$No_of_Votes)
df %>% 
  filter(Released_Year=="2020") %>% 
  filter(No_of_Votes>mean(No_of_Votes),No_of_Votes<374161) %>% 
  group_by(Released_Year) %>% 
  summarise(Meta_score = mean(Meta_score, na.rm=TRUE))  


```

8. (Optional - Challenging). The current `Runtime` variable is not a numeric. Use the `str_remove` function to remove ` min` from the variables then use `as.numeric` to convert the variable to numeric. Calculate the average running time in the 2010s. Calculate the correlation between running time and rating (adding `use="complete.obs"` in the `cor` function to ignore the missing values).  
```{r}


df$Runtime = str_remove(df$Runtime, " min")
df$Runtime = as.numeric(df$Runtime) 
df %>% 
  filter(Released_Year>'2009',Released_Year<'2020') %>% 
  summarise(Mean_runtime = mean(Runtime))


```

9. We can use `select_if` to select columns satisfying a condition and use `summarise_if` to do calculation on columns satisfying a condition.  Try the follows to understand these functions. 

```{r,eval=FALSE}
# Select only character columns
df %>% select_if(is.character)
```


```{r,eval=FALSE}
# Calculate the median of all numeric columns
df %>% summarise_if(is.numeric, mean, na.rm=TRUE)
```


10. Implement the follows functions or combos. Drawing a comment or summary from each calculation. The codes in this question should be different from the codes used in other questions.

- select
```{r}

df %>% 
  select(Released_Year,No_of_Votes,Gross) %>% 
  select_if(is_double)


```

- filter
```{r}

df %>% 
  select(IMDB_Rating,Meta_score,Gross) %>% 
  filter(IMDB_Rating>9 | Meta_score > 90 | Gross > 80000000  )

```

- mutate
```{r}

df %>% 
  mutate(Gross_minus_Votes = Gross - No_of_Votes) %>% 
  select(Gross_minus_Votes)


```

- summarise
```{r}

df %>% 
  filter(Released_Year=="2018") %>% 
  filter(IMDB_Rating>8.2) %>% 
  summarise(Min_Gross = min(Gross, na.rm=TRUE))  

```

- arrange
```{r}

df %>% 
  filter(Gross>100000000) %>% 
  group_by(Director) %>% 
  summarise(Mean_meta = mean(Meta_score, na.rm=TRUE)) %>% 
  arrange(-Mean_meta)

```

- count
```{r}

df %>% count(Released_Year)
  

```

- count + arrange
```{r}

df %>% count(Genre) %>% 
  arrange(-n)

```

- filter + count + arrange
```{r}

df %>% 
  filter(Star1=="Tom Hanks") %>% 
  count(Star2) %>% 
  arrange(-n)

```

- group_by + summarise
```{r}

df %>% 
  group_by(Director) %>% 
  summarise(Sum_Gross = sum(Gross)) %>% 
  arrange(-Sum_Gross)

```

- filter + group_by + summarise
```{r}

df %>% 
  filter(Released_Year=="2019") %>% 
  group_by(Star1) %>% 
  summarise(Sum_Gross = sum(Gross)) %>% 
  arrange(-Sum_Gross)

```

- filter + group_by + summarise + arrange
```{r}

df %>% 
  filter(Meta_score==100) %>% 
  group_by(Director) %>% 
  summarise(Sum_Gross = sum(Gross)) %>% 
  arrange(Sum_Gross)

```
